Hysteresis in human computation:\\ how one task affects another

### 2015-03-23 docs/publications/2014.09_SCIENCE/delete/ban.tex
added all missing files

Here we describe a series of experiments, using the Amazon Mechanical
Turk (MTurk) microtask platform, in which we studied intertask effects
for image-labeling tasks.  Image-labeling tasks are among the most 
common kinds of microtasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, 
and it seems likely that this will remain the case in the long term because
they provide a source of ground truth in computer vision research 
\cite{5543192}.  
During the experiments, workers were presented with a series of images (one at
a time), and had to provide five descriptive words (\textit{labels}) for each.
We varied first five images (hereafter the \textit{initial tasks})
and observed the effect this had on the workers' labels for the last five
(\textit{test tasks}).

The initial tasks were chosen to embed a certain \textit{prime concept}.
In our first experiment, our prime concepts were \textit{food} and (non-food) 
\textit{objects}.  In our second experiment, the prime concepts were 
\textit{food} and \textit{culture}.  The test tasks for the experiments used 
images that combined both prime concepts used in the initial tasks. Workers 
participated in only one treatment of one experiment, and were randomly 
assigned initial tasks based on one of the 
prime concepts.  From the perspective of the worker, we made no distinction 
between the initial tasks and test tasks.

As a point of comparison, we included treatments in which we induced
\textit{framing}.
In the framing treatments, before workers begin the tasks, we told them that 
the purpose of study was to understand the visual perception of 
``food and ingredients'' in one case, or the 
perception of ``objects and tools'' in another.  We varied the language of 
the framing text (hereafter simply the \textit{frame}), and experimented with 
requiring the worker to echo the frame using a combo-box input.

We found that intertask effects were stronger than framing, unless the worker
was required to echo the frame, in which case the effects of both exposures were on par.  
Our results demonstrate that initial tasks can change 
workers' focus in later tasks.  Using the wordnet corpus, we analyzed 
the effects on the vocabulary used by workers when 
labeling test tasks. 
Exposure to both framing and initial tasks significantly altered the 
diversity of vocabulary that workers used when refering to the prime concept,
as well as the degree to which workers used specialized (rather than generic) 
words.

@@

\paragraph*{Measuring computational hysteresis.}
Before describing the details of our experiments, we explain some 
choices made in measuring intertask and framing effects 
(hereafter collectively \textit{exposure effects}).  
In trying to establish a mesure of exposure effects, it might be tempting to 
invoke the notion of a ``neutral frame of mind'' to serve as a baseline. 
But such a concept seems both difficult to define, and unlikely to arise.
Therefore, we always consider the effect of one exposure 
\textit{relative} to another.  We organized our experiments 
into pairs of treatments, between which it is possible to make meaningful 
comparisons.

Our goal in measuring exposure effects is to show that they can actually 
alter the course of algorithms.
We could, by means of a $\chi^2$ test, reject the null hypothesis that
differently exposed HPUs have the same distribution of outputs.
However, this alone would not demonstrate a practically significant effect, in 
terms altering the execution path of algorithms.

@@

\paragraph{Dynamics of intertask effects.} It is natural to expect 
intertask effects to attenuate as a worker proceeds through the test tasks.  
In \textit{task1}, we permuted the test tasks, allowing us to 
measure intertask effects as a function of task position, while factoring
out the content-dependant differences between tasks
(see Fig.~\ref{fig:theta}B). 
We found intertask effects were most pronounced for the first test task,
dropping off substantially for the second, but maintaining a 
significant effect through until the fifth test task.  
This shows that the range of intertask effects extends accross at least five 
tasks, and probably substantially more.
This makes intertask effects of greater concern, but as we will show, it also
makes intertask effects of greater potential utility.
In future work, it would be interesting to investigate whether a worker's
susceptibility to intertask effects varies accross a large batch.  For example,
It is conceivable that the first few tasks induce especially strong
or long-lasting effects, in effect ``setting the tone'' for the rest of the batch.

@@

Using the augmented hypernym-hyponym structure of wordnet to identify nouns 
depicting food, we found that, for the experimental pairings in 
\textit{exp1}, the
food-exposed treatments did produce a significantly larger fraction of 
food-related labels (see Fig.~\ref{fig:specificity}A).  While the shift in 
focus with regard to food does help explain the classifier's ability to 
infer worker exposures, it is insufficient to explain the observed performance
(*).
Furthermore, in \textit{exp2:task2} we see a significant but reversed effect
(see Fig.~\ref{fig:specificity}A), which suggests a more complicated 
relationship between exposure and label usage.
Given the difference between the food-culture and food-object conceptual axes, 
it should perhaps not be surprising that the effect of initial task exposure 
is different.  It would appear that we must be prepared to accomodate multiple 
countervailing factors influencing the number of references to the prime 
concept (food).  We will come back to this point momentarily.

@@

In a crowdsourcing task, considerable effort must be given to shaping the 
vocabulary with which a worker engages a task.  To draw an example from the 
domain of citizen science, in galaxy zoo, workers catalog images of galaxies 
in terms of \textit{smoothness}, and the presence of \textit{rings}, 
\textit{dust lanes}, \textit{digital disturbances}, and so on.  Shaping this 
vocabulary is crucial to the reliability of results.  The interface of
galaxy zoo helps by contstraining and prompting the user with this basic 
vocabulary.  But this would not appear to be sufficient, since
a set of positive and negative training examples are used, together with
peer-to-peer and peer-to-expert exchange in discussion areas, presumably to
further refine and standardize the vocabulary (\#).

@@

\subsection*{Assessing task similarity}
\begin{table}
\centering
\begin{tabular}{ l  s s s s}

\toprule    
Image set   
& \multicolumn{1}{c}{Ambig.} 
& \multicolumn{1}{c}{Cultural} 
& \multicolumn{1}{c}{Ingr.}
& \multicolumn{1}{c}{Test} \\

\midrule

Ambiguous  & 1 & 0.0418 & 0.142 & 0.167 \\

Cultural  & 0.0418  & 1 & 0.0347 & 0.0561 \\

Ingredients  & 0.142  & 0.0347 & 1 & 0.110 \\

Test & 0.167  & 0.0561 & 0.110 & 1
\\
\bottomrule

\end{tabular}
\caption{\footnotesize{
Pairwise similarities of each image set based on the labels attributed to them (see \textbf{Eq. 4}).
}}
\label{table:2}
\end{table}

One weakness of the above analysis is the reliance on the informal notion
of the similarity of one set of images to another. The 
characterization of image content is a deeply complex issue that has been 
approached by many disciplines \cite{panofsky1939studies,shatford1986analyzing,Tversky1977327,Jaimes20002}. \textit{Perceptual similarity} is a 
particularly recalcitrant concept.   But to formalize our claim, we seek a 
measure of similarity between two sets of images, 
\textit{with respect to the labeling task}.  This can be operationalized
in a straightforward way, by using the similarity of the labels given to images
as a proxy for the similarity of the images.  Thus, to measure the 
similarity between two sets of images, X and Y , we computed the Jacquard 
index between the sets of labels attributed to them:
$$
	\mathrm{Sim}(X,Y) = \frac{L(X) \cap L(Y)}{L(X) \cup L(Y)}
$$
where $L(X)$ denotes the set of labels attributed to images in $X$.

The pairwise similarities of the image sets are presented in Table 2. This 
shows that the workers that yielded higher specificity labels were from 
treatments whose initial tasks were more similar to the test tasks (c.f.
\textbf{Fig 3}).

Such a phenomenon would be consistent with the psychological mechanism known as
\textit{negative priming}.

>>>>

Here we describe a series of experiments, using the Amazon Mechanical
Turk (MTurk) microtask platform (\#), in which we studied intertask effects
for image-labeling tasks.  Image-labeling tasks are among the most 
common kinds of microtasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, 
and it seems likely that this will remain the case in the long term because
they provide a source of ground truth in computer vision research 
\cite{5543192}.  
During the experiments, workers were presented with a series of images (one at
a time), and had to provide five descriptive words (\textit{labels}) for each.
We varied first five images (hereafter the \textit{initial tasks})
and observed the effect this had on the workers' labels for the last five
(\textit{test tasks}).

The initial tasks were chosen to embed a certain \textit{prime concept}.
Our first experiment compares the prime concepts \textit{food} and (non-food) 
\textit{objects}, while our second experiment compares \textit{food} and 
\textit{culture}.  For both experiments, The test tasks used images that
combined both prime concepts used for the initial tasks \td{reword}.  Workers participated 
in only one treatment of one experiment, and were randomly assigned initial 
tasks based on one of the prime concepts.  From the perspective of the worker, 
we made no distinction between the initial tasks and test tasks.

As a point of comparison, we included treatments in which we induced
\textit{framing}. \td{what is framing?}
In the framing treatments, before workers begin the tasks, we told them that 
the purpose of study was to understand the visual perception of 
``food and ingredients'' in one case, or the 
perception of ``objects and tools'' in another.  We varied the language of 
the framing text (hereafter simply the \textit{frame}), and for some treatments
we required the worker to echo the frame using a combo-box input.

We found that intertask effects were stronger than framing, unless the worker
was required to echo the frame, in which case the effects of both exposures were on par.  
Our results demonstrate that initial tasks can change 
workers' focus in later tasks.  Using the wordnet corpus, we analyzed 
the effects on the vocabulary used by workers when 
labeling test tasks. 
Exposure to both framing and initial tasks significantly altered the 
diversity of vocabulary that workers used when referring to the prime concept,
as well as the degree to which workers used specialized (rather than generic) 
words.

@@

\paragraph*{Measuring computational hysteresis.}
Before describing the details of our experiments, we explain some 
choices made in measuring intertask and framing effects 
(hereafter collectively \textit{exposure effects}).  
In trying to establish a measure of exposure effects, it might be tempting to 
invoke the notion of a ``neutral frame of mind'' to serve as a baseline. 
But such a concept seems both difficult to define, and unlikely to arise.
Therefore, we always consider the effect of one exposure 
\textit{relative} to another.  We organized our experiments 
into pairs of treatments, between which it is possible to make meaningful 
comparisons.

Our goal in measuring exposure effects is determine whether they can actually 
alter the course of algorithms.
We could, by means of a $\chi^2$ test, reject the null hypothesis that
differently exposed HPUs have the same distribution of outputs.
However, this alone would not demonstrate a practically significant effect, in 
terms altering the execution path of algorithms.

@@

\paragraph{Dynamics of intertask effects.} It is natural to expect 
intertask effects to attenuate as a worker proceeds through the test tasks.  
In \textit{task1}, we permuted the test tasks, allowing us to 
measure intertask effects as a function of task position, while factoring
out the content-dependent differences between tasks
(see Fig.~\ref{fig:theta}B). 
We found intertask effects were most pronounced for the first test task,
dropping off substantially for the second, but maintaining a 
significant effect through until the fifth test task.  
This shows that the range of intertask effects extends across at least five 
tasks, and probably substantially more.
This makes intertask effects of greater concern, but as we will show, it also
makes intertask effects of greater potential utility.
In future work, it would be interesting to investigate whether a worker's
susceptibility to intertask effects varies across a large batch.  For example,
It is conceivable that the first few tasks induce especially strong
or long-lasting effects, in effect ``setting the tone'' for the rest of the batch.

@@

Using the augmented hypernym-hyponym structure of wordnet to identify nouns 
depicting food, we found that, for the experimental pairings in 
\textit{exp1}, the
food-exposed treatments did produce a significantly larger fraction of 
food-related labels (see Fig.~\ref{fig:specificity}A).  While the shift in 
focus with regard to food does help explain the classifier's ability to 
infer worker exposures, it is insufficient to explain the observed performance
(*).
Furthermore, in \textit{exp2:task2} we see a significant but reversed effect
(see Fig.~\ref{fig:specificity}A), which suggests a more complicated 
relationship between exposure and label usage.
Given the difference between the food-culture and food-object conceptual axes, 
it should perhaps not be surprising that the effect of initial task exposure 
is different.  It would appear that we must be prepared to accommodate 
multiple 
countervailing factors influencing the number of references to the prime 
concept (food).  We will come back to this point momentarily.

@@

In a crowdsourcing task, considerable effort must be given to shaping the 
vocabulary with which a worker engages a task.  To draw an example from the 
domain of citizen science, in galaxy zoo, workers catalog images of galaxies 
in terms of \textit{smoothness}, and the presence of \textit{rings}, 
\textit{dust lanes}, \textit{digital disturbances}, and so on.  Shaping this 
vocabulary is crucial to the reliability of results.  The interface of
galaxy zoo helps by constraining and prompting the user with this basic 
vocabulary.  But this would not appear to be sufficient, since
a set of positive and negative training examples are used, together with
peer-to-peer and peer-to-expert exchange in discussion areas, presumably to
further refine and standardize the vocabulary (\#).

@@

\subsection*{S1: Experimental materials}
The tasks for both experiments were presented as a series of slides.  In
both cases, the first slide consisted of a brief set of instructions, followed
by the frame if shown (only some treatments involved a frame), see 
Fig. S\ref{fig:hit_preamble}.  Following the instructions and prime, a set
of initial tasks was shown (except for the treatments in \textit{frame2}
and \textit{echo}) followed by (in all cases) a set of test tasks.  In
\textit{exp1}, two kinds of frame were used, one simple frame, and one where
the worker was required to echo the frame using a combo-box input.  The
instructions, both kinds of frame, and an example task slide, are shown 
in Fig. S1.
\begin{figure}
	\includegraphics[scale=0.8]{figs/tasks.pdf}
	\label{fig:hit_preamble}
	\caption{Examples of A) instructions; B) a frame, as shown in 
		\textit{frame1:obj}, and similar to those shown in 
		\textit{frame1:food} and \textit{frame2:food}, and 
		\textit{frame2:cult};  
		C) an echoed frame, as used in \textit{echo:obj}, and similar to that
		used in \textit{echo:food}; and D) an 
		example of an image labeling task.
	}
\end{figure}

All of the initial tasks had the format shown in Fig. S1D.  The images used
in the initial tasks for \textit{task1:food} and \textit{task1:obj} are
shown in Fig. S2 and S3, and the images used in the test tasks for 
\textit{exp1} are shown in Fig. S4.  The images used in the initial tasks
for \textit{task2:food} and \textit{task2:cult} are shown in Fig. S5 and
Fig. S6 respectively, and the images used in the test tasks for \textit{exp2} 
are shown in Fig. S7.

\begin{figure}
	\includegraphics{figs/task1-food.pdf}
	\label{fig:task1:food}
	\caption{
		Figure S2: Images used in the initial tasks for 
		\textit{task1:food}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task1-obj.pdf}
	\label{fig:task1:obj}
	\caption{
		Figure S3: Images used in the initial tasks for 
		\textit{task1:obj}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task1-test.pdf}
	\label{fig:task1:test}
	\caption{
		Figure S2: Images used in the test tasks for 
		\textit{exp1}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task2-food.pdf}
	\label{fig:task2:food}
	\caption{
		Figure S2: Images used in the initial tasks for 
		\textit{task2:food}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task2-cult.pdf}
	\label{fig:task2:cult}
	\caption{
		Figure S2: Images used in the initial tasks for 
		\textit{task2:cult}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task2-test.pdf}
	\label{fig:task2:test}
	\caption{
		Figure S2: Images used in the test tasks for 
		\textit{exp2}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\subsection*{S2: Access to raw data and code}
The raw data and source code needed to reproduce the analyses 
shown here and in the main text can be downloaded from 
http://networkdynamics.org/publications/public/priming2014.09.

\subsection*{S3: Additional results not shown in the main text}
		the top 5 most suppressed and most activated terms between treatment
		pairs.

\begin{table}
	\centering
	\setlength{\tabcolsep}{10pt}
	\begin{tabular}{ c c c c c }

		\setlength{\tabcolsep}{4pt}
		\begin{tabular}{ r | c }
		\toprule
		\multicolumn{2}{c}{\textit{task1}} \\
		\toprule
		coffee & 38 \\
		meal & 34 \\
		cheese & 34 \\
		apple & 32 \\
		dessert & 21 \\
		cup & -30 \\
		glass & -45 \\
		table & -70 \\
		candle & -74 \\
		food & -80 \\
		\bottomrule
		\end{tabular}

&

		\setlength{\tabcolsep}{4pt}
		\begin{tabular}{ r | c }
		\toprule
		\multicolumn{2}{c}{\textit{frame1}} \\
		\toprule
		bread & 18 \\
		wine & 18 \\
		cheese & 16 \\
		apple & 14 \\
		oil & 12 \\
		table & -9 \\
		meal & -10 \\
		candle & -12 \\
		dinner & -13 \\
		food & -32 \\
		\bottomrule
		\end{tabular}

&

		\setlength{\tabcolsep}{4pt}
		\begin{tabular}{ r | c }
		\toprule
		\multicolumn{2}{c}{\textit{echo}} \\
		\toprule
		apple & 24 \\
		cheese & 23 \\
		wine & 15 \\
		coffee & 14 \\
		oil & 7 \\
		knife & -24 \\
		dinner & -26 \\
		fork & -27 \\
		candle & -35 \\
		food & -55 \\
		\bottomrule
		\end{tabular}

&

		\setlength{\tabcolsep}{4pt}
		\begin{tabular}{ r | c }
		\toprule
		\multicolumn{2}{c}{\textit{task2}} \\
		\toprule
		spicy & 26 \\
		sauce & 17 \\
		indian & 15 \\
		buffet & 14 \\
		exotic & 12 \\
		festival & -11 \\
		offering & -12 \\
		statue & -15 \\
		india & -20 \\
		food & -56 \\
		\bottomrule
		\end{tabular}

&

		\setlength{\tabcolsep}{4pt}
		\begin{tabular}{ r | c }
		\toprule
		\multicolumn{2}{c}{\textit{frame2}} \\
		\toprule
		indian & 11 \\
		banquet & 8 \\
		spicy & 7 \\
		asian & 6 \\
		variety & 6 \\
		delicious & -6 \\
		meat & -7 \\
		festival & -7 \\
		spice & -7 \\
		food & -9 \\
		\bottomrule
		\end{tabular}

	\end{tabular}
	\caption{blah}
\end{table}

\subsection*{S4: Data preprocessing}
	\paragraph{Splitting, Lemmatization, removal of stop-words, and 
		addition of position tags.}

	Before performing any analysis on the labels that workers provided, we
	performed a series of preprocessing steps.  
	\td{does this reflect the actual order in which the steps were done?}
	First, words wore lematized 
	using the
	\td{whatzit} lemmatizer.  This lemmatizer is very mild, tending only to 
	change plural words to their singular form.  Common
	words such as ``the'', ``to'', or ``with'', were found using the
	the \td{whatzit} stop-word list and removed.  Then, misspelled words were
	automatically corrected using a spelling correction algorithm described 
	below.  Finally, labels that contained
	multiple words (separated by spaces or punctuation) were split, with
	punctuation removed, and the separated words were treated as distinct 
	features in subsequent analysis.

	For the purpose of training and testing a naive Bayes classifier, we 
	performed an additional preprocessing step.  We tagged words with the
	position in which they had been entered (i.e. which of the five text 
	inputs in the task interface) as well as the test tasks in which the word 
	had been provided.
	So, if the word ``wine'' was entered into the second text-input for 
	the third test-task, after preprocessing, the feature ``3\_2\_wine'' would
	appear.  Prepending the task number onto words was simply a means to 
	retain correct attribution, which itself was essential since providing a
	word during one task is not equivalent to providing the same word during 
	another task.

	\paragraph{Spelling correction.}  
	Spelling correction was performed using an algorithm that first detected
	if a word was likely to be misspelled, then generated a set of candidate 
	corrections, and chose the best candidate based on a scoring mechanism.

	A word was considered misspelled if it was not contained in the 
	\textit{legal set}, which was formed by the union of
	the wordnet corpus, the stop-word list, and a set of words seen while 
	crawling the world food section of the allrecipes.com website.  We
	describe the crawling of the allrecipes.com website in a section below.

	To correct misspellings, we first produced all possible modified forms 
	that could be obtained applying one or two edits.  An edit consisted of 
	adding or removing a letter, changing one letter into another, or 
	swapping the positions of two adjacent letters.  For the purpose of these 
	edits, spaces were treated as any other letter.

	The candidates produced by these edits which were in the legal set were
	then ranked based on a scoring mechanism, and the highest scoring word
	was chosen as the correction.  A candidate $w$'s score, $s_w$, was 
	calculated according to the following formula:
	\begin{equation}
		s_w = (f_w + 1) * p_1 * p_2,
	\end{equation}
	where $f_w$ is the frequency with wich the word occured (correctly 
	spelled)
	within the given task, $p_1$ was a penalty for the first edit, and
	$p_2$ was a penalty for the second edit.  If the word was made using only
	one edit, then $p_2 = 1$.  Any edit that did not involve adding a space
	(i.e. separating a word) incured a penalty of 0.5, while the addition of
	a space incurred a penalty of 0.1.  Word-separation edits were more 
	strongly penalized because it is often possible to split a series of 
	letters into many short two- or three-letter words, which otherwise leads 
	to many erroneous corrections.  We found that setting the penalty for
	word splitting at 0.1 worked well by testing on words taken from
	initial tasks.

	After all of our analyses had been performed, we checked the accuracy of 
	the spelling correction algorithm using three human coders.  The coders
	were shown a set of 
	500 words that were randomly sampled from the labels attirubuted during 
	test tasks (50 for each test task).  Before sampling, the labels from all
	experimental treatments, for the given test task, were pooled together.
	The coders did not know which treatment any given word came from.  
	In addition to the words, coders were shown the spelling correction 
	produced by the algorithm, as well as the image from the test task.
	The coders were asked to identify any misspelled words which had not 
	been corrected, as well as any corrections that appeared to be erroneous,
	by indicating their own correction.

	Based on this exercise, we found that, according to the human coders,
	17\% of the words were misspelled, but that after applying the 
	automated spelling correction, only 3.2\% of words were misspelled.

	\paragraph{Crawling the world food section of allrecipes.com.}
	The website allrecipes.com was accessed on 3-4 November 2014 using 
	automated scripts.  A total of 2642 recipe listings and 15621 recipe
	instruction pages were crawled.  Recipe listings were pages that provided 
	lists of recipes, and contained a title and short description for each.  
	The recipe instruction pages had lists of ingredients and preparation 
	instructions.  All of the words found in recipe titles, short
	descriptions, ingredients lists, and preparation instructions were
	collected, and saved as an auxiliary set to aument the \textit{legal set}
	used in spelling correction.

\subsection*{S5: Identification of food-related terms}
The wordnet corpus is composed of synsets, which are particular senses 
(meanings) and a set of word forms bearing that meaning. A given word form, 
such as ``ring'' has multiple senses (``to ring a bell'', ``a wedding ring''), 
and so can be part of many synsets.  Since the wordnet corpus was designed
to carry semantic information, the synset is the basic organizing element
of wordnet.

We chose the synsets \textit{food.n.01} and \textit{food.n.02} to act as roots
in defining which words should be considered food-related.  Any word form
belonging to a synset which was a hyponym of one of the two root synsets
identified above was considered to be a reference to food.  Here, as in the
main text, when we say hyponym, we mean either a direct hyponym, or a hyponym
of a hyponym, and so on.

This is a stringent notion of a word being a food reference.  It roughly
corresponds to whether or not a given thing is reasonably considered 
consumable.  So, ``orange'' is food, but ``salty'' is not.  Although ``salty''
would usually refer to something edible, it is not itself an edible thing.
On the other hand, ``salt'' would be considered a reference to food under 
this operationalization.

The subset of the wordnet corpus induced by the hyponyms of \textit{food.n.01}
and \textit{food.n.02} provide an extensive set of \td{X} words.  Manual
testing appeared to show that the coverage was very extensive, except in 
some cases references to ethnic foods were missing.  Since, especially in
\textit{exp2} we had images containing many ethnic foods, we decided to 
augment the wordnet corpus with words learned by crawling the world food
section of the allrecipes.com website.

After collecting a set of words using an automated script 
(\td{described above}), we collected the set of words that had been used by
workers and which had been seen during crawling of the allrecipes.com website,
but which were not included in wordnet.  We graphted these extra words into
the wordnet hyponym-hypernym graph manually, which ensured that any references
to ethnic words learned from crawling would be properly classified as 
food references.

\paragraph{Validation of the detection of food references.}
To determine whether our extended version of wordnet provided a good approach
to detecting food-related words, we sampled 500 words from the set of labels
produced by workers and had three human coders manually decide if they were 
food-related or not.  The method for randomly sampling the 500 words was 
described in above in \td{section number}.

The coders were instructed to consider whether a word was a noun signifying
an edible item.  The guiding principle was, for the coder to ask herself,
``can I eat this X'', where X is the word she is coding.  The coders were
shown the used in the task from which the word came, to help resolve ambiguity
in the sense of the word that had been given.  Coders were also shown the 
spelling correction (if any) that the spell-correction algorithm had made,
to help interpret mispelled words \td{we describe the spelling correction 
algorithm in section X}.

We evaluated the food detection algorithm in two ways.  First, we measured
the inter-rater reliability between the three human coders and the algorithm
(treating the algorithm just like any other coder).  Second, adopting the 
majority code given by the human coders, we determined the accuracy of the
food-detection algorithm.  Data sumarizing this validation process are given
in table S1.

\begin{table}
\centering
\setlength{\tabcolsep}{12pt}
\begin{tabular}{ r | c }
\toprule    
\# terms coded & 500 \\
\# food references & 130 \\
\# correct machine codes & 440 \\
inter-rater reliability & 82.4\% \\
human-machine code agreement & 88\% \\
%machine recall & 80 \% \\
%machine precision & 75.4 \% \\
%machine code F1 & 0.78 \\
\bottomrule
\end{tabular}
\caption{\footnotesize{
	\td{remove the automatic numbering}. Table S1. Results for the validation 
	of the automatic detection of food references, by comparison to
	human-generated coding on randomly sampled words.  In calculating the 
	human-machine code agreement, the majority human code was used.
	Inter-rater reliability was calculated as Krippendorff's alpha.
}}
\end{table}

The results show that good inter-rater reliability was achieved during the
validation (82.4\%).  Looking at the agreement between the 
machine codes and the majority human-generated codes, we find they agree
88\% of the time.  This means that the concept used to instruct human coders
provides a good approximation to what the wordnet-based machine codes actually
represent. In other words, the machine codes correspond closely to indicating
which words correspond to nouns that refer to edible things, and which ones
don't, which provides a clear and simple interpretation for the machine
coding of food and non-food words.

\subsection*{S6: Calculation of relative specificity}
In the main text we present results for the relative specificity of labels
produced by two experimental treatments.  In performing this calculation,
we used the wordnet corpus, which has a set of hypernym-hyponym relationships
as described in the main text.  The first step in performing this calculation
was therefore to map the words occuring from (the labels of) a given 
experimental treatment onto the wordnet synsets.  In general, a given word
form can have multiple meanings, and therefore maps onto multiple wordnet 
synsets.

The result of mapping the words from two treatments onto the wordnet 
hypernym-hyponym graph two sets of counts, one set per treatment, 
indicating the number of times a word from a given synset occured among the
labels of the treatment.  Then, for every synset in one treatment, we looked
at the number of synsets in the other treatment that were more generic
(reachable by following hypernym relations) less the number of words that
were more specific (reachable by following hyponym relations).  This quantity
tallied for all words in the original synset gives a (non-normalied) measure
of the degree to which words from the first treatment are more specific
than words from the second treatment, overall.  We then normalized this 
quantity by the total number of possible comparisons between the words of
one treatment to those of the other, where two words are considered comparable
only if one word is the hypernym or hyponym of the other.  Hence, ``statue'' 
and ``bread'' are not comparable, but ``pumpernickel'' and ``bread'' are.

This calculation can be summarized by the following equation:
\begin{equation}
	S(P,Q) = \frac{
		\sum_{w\in P}\sum_{v\in Q} \left(
			\mathbf{1}_{[w>v]} - \mathbf{1}_{[v>w]} \right)
	}{
		\sum_{w\in P}\sum_{v\in Q} \left(
			\mathbf{1}_{[w>v]} + \mathbf{1}_{[v>w]} \right)
	},
\end{equation}
where $P$ and $Q$ are sets of synsets, and $\mathbf{1}_{[w>v]}$ evaluates
to 1 if synset $w$ is more specific than (i.e. is a hyponym of) synset $v$.
This measure counts the excess of cases where synsets from $P$ are more
specific then synsets from $Q$, as a fraction of all comparable synset pairs.

In computing this quantity between two treatments, we first computed it 
using the labels attributed to a particular test task, and then
averaged the result obtained in this way for each of the five test tasks.

\subsection*{S7: Analysis of variance}
	\paragraph{Variance in the average fraction food-related labels}
	In the main text, Fig.~\ref{fig:specificity}A shows the difference in 
	the fraction of food related labels used by workers in from different 
	treatments.  The error bars show the 95\% confidence intervals around 
	the estimated difference of means.  It was computed from the
	observed sample standard deviation in the fraction of food labels given
	by individual workers.

	\paragraph{Variance in food vocabulary size.}
	\td{Not sure how to calculate this yet.}

	\paragraph{Variance in food specificity.}
	\td{Need to calculate this using bootstrapping.}

\subsection*{S8: Measuring computational hysteresis}
	\paragraph{Using cross-validation accuracy to lower-bound L1-distance.}
	In order for Eq.~\ref{l1} to be applicable, it is necessary to obtain
	an unbiased estimate of a classifier's performance.  One well-known
	pitfall in measuring this error is to underestimate it due to 
	over-fitting of the data used.  As an example, if one were to try 
	a large number of classifiers, eventually one would find one that
	classifies the data perfectly, even though it might non generalize to
	new data.

	It is customary to use part of a data set to optimize hyperparameters of
	a model, and then test the model on as-yet unseen data.  In our case,
	don't have enough samples in each treatment to do this.  Instead, for 
	this reason, we make principled decisions for all of the options used
	in our classifier, and then test it's accuracy using cross validation,
	with no optimization performed on the data.  It was cruicial for 
	the statistical validity of our approach that we test our classifier on 
	each datum only once, and report these results without any optimizations
	of the classifier.

	In performing the classification we had various options, and since we
	could not optimize these choices, we made principled decisions based on
	prior knowledge from text classification applications.  We describe the
	options we faced and the choices we made below.

	\textit{whether to correct misspelled words.}  Because we used free-text
	entry for the labels, there was bound to be many spelling mistakes.
	This would tend to reduce the accuracy of a classifier, because the 
	entries ``bread'' and ``braed'' would be regarded as completely different
	features, even though they clearly correspond to the same 
	\textit{intended word}.  While no algorithm can determine a worker's 
	intentions, it is possible to check words against a corpus, and, in the
	case of a misspelled word, look for plausible corrections based on 
	edit distance.

	\textit{Lemmatization.}  In some cases it is helpful to simplify words 
	before providing them to the classifier.  For example, the words 
	``berries'' and ``berry'' only differ in number, but would normally be
	treated as completely different features.  Lemmatization ensures that
	nouns are in their sigular form, which helps to eliminate variations in
	labels that increase sparsity without conveying discriminative 
	information.

	\textit{Removal of stop-words.}  Certain very common words such as `the',
	``to'', and ``it'' are extremely common but don't help discriminate a 
	worker's exposure.  Removing these words is a common practice because
	they otherwise introduce noise into the classifier learning process.

	\textit{Whether to split multiple words.}  Although workers were 
	instructed to put one word per text input, many put multiple words.
	Theoretically the original multi-word entry might bear additional 
	information that could be useful in discerning a worker's exposure. 
	In practice, it would be necessary for that word combination to recur,
	which is unlikely.  Therefore we chose to split multi-word entries,
	using the individual words as features.  This meant that the outputs of
	some workers generated more features than those of others.

	\textit{Preserving the location of labels.}  When workers completed image
	labeling tasks, they had to fill five text inputs, which were stacked
	vertically.  Presumably, they began with the top input and proceeded 
	downwards.  It seems likely that exposure effects might influence the
	\textit{order} in which workers apply labels, so when processing the 
	labels we prepended an integer to indicate into which text input the 
	label been input.  This meant that the classifier could distinguish
	``bread'' entered into the first text input from ``bread'' entered into
	the second.  A drawback is that this increases sparsity, preventing the
	classifier from accounting for the fact that different though they are,
	these entries should be treated similarly.  For this reason, we also
	provided features to the classifier which did not encode the particular
	input that had been used.  Thus, two workers that entered ``bread''
	into different inputs would generate both a feature in common indicating
	the use of the word ``bread'', as well as distinct features, indicating
	entry of the word ``bread'' \textit{in a particular input}.

	\textit{Preserving the connection between tasks and labels.}  In many
	cases a particular label would appear in many tasks.  For example, the
	word ``food'' appears in all test tasks.  It stands to reason that
	exposure effects play out differently for different tasks, and that
	the attribution of a particular label during one task provides different
	information from the attribution of the same label during a different 
	task.  So, as we did in the case of input location, we prepended labels
	with the identity of the task in which they were added, so that the 
	two otherwise identical labels added for different tasks were treated
	by the classifier as unique features.  Unlike in the case of input 
	position attribution, we didn't include raw labels unadorned with 
	the task number.  Our rationale for this is that the task defines the 
	context in which a label is applied, and so the elicitation of a label 
	during one task represents a distinct output from the elicitaiton of the 
	same label during a different task.

\paragraph{Selection of the naive Bayes classifier.}
A great deal of work has been done in the application of classifiers to
text documents, which provides the bacground for our choice of classifier.
The Support Vector Machine (SVM) approach is known to perform well, better in 
general than naive Bayes.  However, three considerations drove us to use
a naive Bayes classifier in favor of SVM.

First, we used 119 examples (workers) per treatment, but the number of 
features
for each worker (i.e. the number of distinct labels, or modifications 
thereof produced in preprocessing), was generally in the thousands.  
This tends to
be problematic for many classifier algorithms, including SVM, because it
leads to very high variance models, which manifests as overfitting and poor
performance.  To deal with this, some approach to culling features is 
generally used, but this raises the issue of how to decide the culling in 
a principled manner, generally adding further hyperparameters which could
be optimized.  However, one of the strengths of the naive Bayes classifier
is that it's performance is not generally degraded from the inclusion of 
large numbers of features, even when the number of examples is comparatively
small.  We expected this feature of the naive Bayes classifier to be 
advantageous in our case.

Second, the standard implementation of the naive Bayes classifier has no 
hyperparameters to optimize.  This allows us to forgo the common practice 
of holding out a test set, and instead use cross-validation as a method to
estimate the true classification error, \textit{since we did not perform
any kind of model selection or optimization whatsoever using our data}.
This point is crucial, because if we did optimize our classifier during 
cross-validation, then the validation error would no longer provide an 
unbiased estimate of the true classification error, but would instead lead
to us over-reporting exposure effects.

Third, the naive Bayes algorithm is based on an assumption that the 
probability of observing one feature is independant from the observation
of any other, \textit{when conditioned on the class}.  This
\textit{conditional independance} assumption is generally made for
pragmatic rather than theoretical reasons, and is rarely satisfied.  However,
if it \textit{is} satisfied, then the naive Bayes classifier is optimal.
In our case, conditional independence means that, provided we know the 
exposure of a worker, knowing one label they provided to an image does
not in any way help us guess any other label for that image.  In most
text documents, the existence of topics and language structure certainly
vioalates the conditional independance of words.  But in our case, such 
effects are less likely to carry from one text input to the next, and in the
small space provided by five inputs, any overarching topics are likely to
be in relation to the image itself, upon which we are conditioning.  In other
words, we believe that, in our setup, conditional independence is better 
satisfied than in most text classification contexts.  That being the case,
one would not expect much improvement in using a more sophisticated algorithm,
such as SVM.

\paragraph{Comparison to other classifiers.}
